
```{r }
---
title: "PA1_template.Rmd"
author: "MJM Beuken"
date: "1 oktober 2017"
output: html_document
---

```
{r, echo=TRUE, results='hide', warning=FALSE, message=FALSE}

library(ggplot2)
library(dplyr)
library(lubridate)
```

## Getting and cleaning the data
#### Loading to csv and getting rit of quotation marks
```{r, results='markup', warning=TRUE, message=TRUE}
setwd('./Data science cursus coursera/course5week2')
data <- read.csv('./activity.csv', quote="\"")

head(data)
```

#### Transform data
```{r}
activity <- na.omit(data)

activity.day <- summarize(group_by(activity, date), steps=sum(steps))
```

-----


##       What is mean total number of steps taken per day?
### 1.   Calculate the total number of steps taken per day
```{r}
Sumsteps <- aggregate(data$steps, by=list(data$date), sum)

names(Sumsteps)[1] ="Date"
names(Sumsteps)[2] ="Total"

head(Sumsteps,15)
```


### 2.   Make a histogram of the total number of steps taken each day
```{r}

ggplot(Sumsteps, aes(x = Total)) + geom_histogram(fill = "blue",col="red", binwidth=1000, center=500) + labs(title = "Steps taken a day", x = "Steps a day", y = "Frequency")

```

### 3.   Calculate and report the mean and median of the total number of steps taken per day
```{r}
summary(activity.day)
```
* Mean: `activity.day`
* Median:  `activity.day`

-----

##       What is the average daily activity pattern?
###      1.      Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
activity.interval <- summarize(group_by(activity, interval), steps=mean(steps))


ggplot(activity.interval, aes(interval, steps)) + geom_line(colour="blue", size=0.3)

```

### 2.      Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
activity.interval[activity.interval$steps == max(activity.interval$steps), ]
```

* Most Steps at: `Interval = 835; average = 206.17`

----

## Imputing missing values
### 1. Calculate and report the total number of missing values in the dataset 
```{r}
nrow(data) - nrow(activity)
```

* Number of missing values: `2304`

### 2. Devise a strategy for filling in all of the missing values in the dataset.
```{r}
names(activity.interval)[2] <- "mean"
activity.filling <- merge(data, activity.interval)

head(activity.filling)
```
### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
activity.filling$steps[is.na(activity.filling$steps)] <- activity.filling$mean[is.na(activity.filling$steps)]

head(activity.filling)
```

### 4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.
```{r}
SumData <- aggregate(data$steps, by=list(data$date), sum)

head(SumData,15)

names(SumData)[1] ="date"
names(SumData)[2] ="total"
head(SumData,15)


ggplot(SumData, aes(x = total)) + geom_histogram(fill = "blue", col="red", binwidth=1000, center=500) + labs(title = "Total Steps a day", x = "Steps", y = "Frequency")


summary(SumData, digits=10)
summary(Sumsteps, digits=10)
```

* Mean (Imputed): `10766.19`
* Median (Imputed):  `10765`

## Are there differences in activity patterns between weekdays and weekends?
### 1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r}
activity$weekday <- weekdays(as.Date(activity$date))
activity$weekend <- as.factor(activity$weekday == "zaterdag" | activity$weekday == "zondag")
levels(activity$weekend) <- c("Weekday", "Weekend")

head(activity, 15)
tail(activity, 15)
```

### 2. Make a panel plot containing a time series plot

```{r}
activity.weekday <- activity[activity$weekend=="Weekday",]
activity.weekend <- activity[activity$weekend=="Weekend",]

activity.interval.weekday <- group_by(activity.weekday, interval)
activity.interval.weekday <- summarize(activity.interval.weekday, steps=mean(steps))

activity.interval.weekday$weekend <- "Weekday"

activity.interval.weekend <- group_by(activity.weekend, interval)
activity.interval.weekend <- summarize(activity.interval.weekend, steps=mean(steps))

activity.interval.weekend$weekend <- "Weekend"

activity.interval <- rbind(activity.interval.weekday, activity.interval.weekend)
activity.interval$weekend <- as.factor(activity.interval$weekend)
plot <- ggplot(activity.interval, aes(interval, steps, color=weekend)) + geom_line(size=0.5) + xlab("Interval") + ylab("Avarage steps")

plot + theme(panel.background = element_rect(fill = 'white'))

```
```

